---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Cross Section Assignment"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Nico Katzke}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Jacques Rossouw"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellebosch, Western Cape, South Africa" # First Author's Affiliation
Email1: "gerardrossouw\\@gmail.com" # First Author's Email address

# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.
# 
# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"
# 
# CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE
# 
# # Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
# Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: TRUE         # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt    # Set fontsize
linestretch: 1.2  # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: FALSE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
---

<!-- First: default preferences for chunk options: -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
        fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H',
        purl = TRUE)

pacman::p_load(tidyverse, dplyr, lubridate, AER, 
 broom, plm, corrplot, psych, ggrepel,
 gridExtra, stargazer, formatR)

fmxdat::source_all("./code")

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->


\newpage

# Introduction


The effects of the covid-19 corona virus has led to global economic strain and
the debate over whether strict lockdown rules were necessary inspite of their 
economic implications is debated by some. 

The problem with identifying the effect that covid policy stringency had on
covid-19 is with the idiosynctratic differences between countries that also 
contributed to the severity of how covid would have affected individuals. 
This suggests that the fixed effects regression technique offers a unique 
opportunity to account for these fixed effects between countries, and to isolate
the effect that policy had on covid. Towards fighting the coronavirus as 
a different flu from the typical cold, the amount of people hospitalised, 
those ending up in ICU, or deaths per unit of time is measured relative to 
the amount of new cases. The typical flu also spreads rapidly but the the
problem is set up as mentioned, to isolate the effect over and above being
a standard flu.

By using the data from @dataset, the data is first cleaned by filtering for 
only the country components, and removing the countries for which data does 
not exist after a certain date.

<!-- Note: All functional code is contained in the appendix -->


# Data Cleaning/Feature Selection

```{r View continents, appendix = T, include=FALSE}

# Checking the date of first observations
start_date()

```


Some transformations are also made to columns to make them more usable for 
regression. The variables that are distributed on a wider range, or scale, 
are also scaled to ensure that the OLS estimation is not biased by this. The
code in the [Appendix] illustrates this.

```{r Fetch, echo=FALSE, include=FALSE, purl=TRUE}

# fetching and cleaning the dataset
world_df <- extract_all() %>% 
    feature_adj_all() %>% 
    experiment_aggregate_week() %>% 
    experiment_trim() %>% 
    relocate(afflicted_rate, .before = reproduction_rate)

world_df
```


## Converting to Cumulative Values


```{r echo=TRUE}
# See appendix
world_df <- world_df %>% scale_bigs_cumsum(.)
```

## Variable Features Scaling

```{r range_var, echo=FALSE}
options(scipen=999)
cols_range(world_df)
options(scipen=0)
```

Thus, want to scale: `new_test, new_vaccinations`

Plotting to see whether there is any irregularity in the distribution of the dependent variable.

```{r afflicted, echo=FALSE}
afflicted_plot(world_df)
```


```{r var_plot, echo=FALSE}
p1 <- cumulative_test_plot(world_df)
p2 <- cumulative_vax_plot(world_df)
grid.arrange(p1, p2, ncol=2)
```


```{r scale_var, echo=TRUE}
world_df <- world_df %>% scale_bigs_scale(.)
```


## Fixed Effects Feature Scaling

Now, to check the scales of the features that remain constant per country:

```{r range_cons, echo=FALSE}
cols_range_constant(world_df)
```

Additional features that need to be scaled are this
- `gdp_per_capita`
- `population_density`
- `cardiovasc_death_rate`

```{r cons_plot, echo=FALSE}

g1 <- cardio_plot(world_df)
g2 <- gdp_plot(world_df)
g3 <- pop_density_plot(world_df)
grid.arrange(g1, g2, g3, nrow=2)
```


```{r scale_cons, echo=TRUE}
world_df <- world_df %>% scale_bigs_constant(.)
```


Now we can check all the descriptive stats for all the columns
```{r descr, echo=FALSE}
options(scipen=999)
world_df %>% cols_range(df = ., constant_features = c("location", "date"))
options(scipen=0)
```

The value of the stringency index is assumed to have a delayed effect on the 
coronavirus, therefore, to account for this, one quarter lagged average 
stringency index value is associated the current with each current period. 

```{r lag}
world_df <- world_df %>%
    group_by(location) %>%
    mutate(across(date, function(x) floor_date(x, unit = "quarters"))) %>% 
    ungroup()
    
quick_df <- world_df %>%
    select(location, date, stringency_index) %>% 
    group_by(location) %>% 
    mutate(across(date, function(x) x %m+% months(3))) %>% 
    filter(date != last(date))

world_df <- quick_df %>% left_join(world_df, by = c("location", "date")) %>% 
    select(-stringency_index.y)
```



# Correlation Analysis


```{r cor_plot, echo=FALSE}
cor_plot1(world_df)$arg$type
```

```{r core_plot2, echo=FALSE, message=FALSE, warning=FALSE}
cor_plot2(world_df)$arg$type
```



```{r include=FALSE}
names(world_df)
```

The components that will form a part of the larger OLS regression is therefore
the:

- stringency index `stringency_index.x`

<!-- - average life expectancy `life_expectancy` -->

- development index `human_development_index`

- proportion of smokers `smokers`

- population over the age of 65 `aged_65_older`

- and hospital beds per 1000 `hosp_beds_1k`

- Availability of hand washing facilities `handwashing_facilities`


Hand washing facilities is added based on intuitive interpretation

# Regressions

## OLS Regression

```{r OLS}
mod_ols_1 <- plm(afflicted_rate ~ stringency_index.x +
   handwashing_facilities + reproduction_rate,
    index = c("location", "date"), data = world_df, 
   model = "pooling")

mod_ols_2 <- plm(afflicted_rate ~ stringency_index.x + smokers
   + handwashing_facilities + aged_65_older + hosp_beds_1k
   + human_development_index + reproduction_rate, 
   data = world_df, 
   index = c("location", "date"), model = "pooling")

mod_1sls <- plm(stringency_index.x ~ gdp_per_capita_log,
                data = world_df, 
                index = c("location", "date"), model = "pooling")

stringency_hat <- fitted.values(mod_1sls)

mod_2sls <- plm(afflicted_rate ~ stringency_hat + smokers
   + handwashing_facilities + aged_65_older + hosp_beds_1k
   + human_development_index, 
   data = world_df, 
   index = c("location", "date"), model = "pooling")

# To get robust standard errors
robustse_ols1 <- sqrt(diag(vcovHC(mod_ols_1, type = "HC1")))
robustse_ols2 <- sqrt(diag(vcovHC(mod_ols_2, type = "HC1")))
robustse_2sls <- sqrt(diag(vcovHC(mod_2sls, type = "HC1")))

```


```{r ols_star, results='asis', message=F, comment='%', tidy=T}

stargazer(mod_ols_1, mod_ols_2, mod_2sls, header = F,
          font.size = "footnotesize", se = list(robustse_ols1,
         robustse_ols2,
         robustse_2sls), column.labels = c("OLS", 
                                             "OLS",
                                             "2SLS"))

```


## Fixed- and Random Effects Regression


```{r FE}

mod_fe_1 <- plm(afflicted_rate ~ stringency_index.x + smokers
   + handwashing_facilities + aged_65_older + hosp_beds_1k
   + human_development_index, 
   data = world_df, 
   index = c("location"), 
  model = "within", effect = "individual")


robustse_fe1 <- sqrt(diag(vcovHC(mod_fe_1, type = "HC1")))

```


```{r RE}
mod_re_1 <- plm(afflicted_rate ~ stringency_index.x + smokers
   + handwashing_facilities + aged_65_older + hosp_beds_1k
   + human_development_index,
                   data = world_df,
                   index = c("location"),
                  model = "random")

robustse_re1 <- sqrt(diag(vcovHC(mod_re_1, type = "HC1")))
```


```{r fe_star, results='asis', message=F, comment='%', tidy=T}
stargazer(mod_fe_1, mod_re_1, header = F,
          font.size = "small", se = list(robustse_re1), column.labels = c("FE", 
                                             "RE"))
```


## Interpretations

### Measurement Error & Bias

It is important to note that the unexpected nature of Covid-19 resulted in
a lot of "on the fly" data collection which can lead to larger measurement 
errors. The reliability of measurement error and false estimates in this case 
is questionable. Furthermore, the fixed- and random effects regressions can
exacerbate this problem. In that case, it would suggest that a pooled OLS 
regression offer some particular benefits.

### Assumptions of OLS:

There exists an inherent endogeneity in this dataset, since there exists some
country specific factors that might have led to lockdowns being more effective
or having more resources to withstand the aderse effects of lockowns. For this
reason, the last OLS model includes a two stage least squares regression model
to mitigate this issue. Additionally, the Random- and Fixed effects models are
also fitted. These can more precisely account for country specific factors 
playing a role.

The Instrumental Variable regression, by method of two stage least squares, 
requires the following assumptions to hold for an unbiased estimate:

Instrument Variable: reproduction rate, such that it affects the afflicted rate
stringency through the stringency index

1. Exclusion Restriction

This requires that the reproduction rate only affects the afflicted rate through
the stringency index.

This assumption is less likely to hold. however, since the assumption is that
the degree of harm that the virus inflicts should reduce even if the 
infectiousness rises due to vaccinations etc. There is still a possibility 
that this might hold. Also, the expection is that if lockdowns are reasonably
effective, that the contagiousness of the virus should not affect the 
afflicted rate per ten thousand individuals in the population.

2. Random Assignment

This requires that there be no difference between the potential outcomes 

### Assumptions of Fixed- and Random Effects



# References

<div id="refs"></div>

# Appendix

```{r get-labels, echo = FALSE}
labs = knitr::all_labels()
labs = setdiff(labs, c("setup","get-labels", "function_scripts",
                       "fe_star", "ols_star", "FE", "RE", "OLS",
                       "cor_plot", "descr", "scale_cons", "cons_plot",
                       "range_cons", "scale_var", "var_plot", "afflicted",
                       "range_var"))
```

```{r all-code, ref.label=labs, eval=FALSE}
```


## Functional Code

```{r function_scripts, echo=FALSE}

afflicted_plot
cardio_plot
cols_range 
cols_range_constant 
cor_plot1
cor_plot2
cumulative_test_plot  
cumulative_vax_plot  
experiment_aggregate_week
experiment_trim 
extract_all
extract_continents  
feature_adj_all 
gdp_plot
pop_density_plot
scale_bigs_constant
scale_bigs_cumsum  
scale_bigs_scale
start_date
```
